Have you forgotten?' Conversion experiences in shopping lists

### 2016-10-26 lists_conversion_experience.rmd
log updates

---
title: "'Have you forgotten?' Conversion experiences in lists"
author: "Adrian Mackenzie"
date: "`r Sys.Date()`"
output:
tufte::tufte_handout: default
citation_package: biber
latex_engine: xelatex
bibliography: ~/ref_bibs/uni.bib
link-citations: yes
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

# 'Have you forgotten?' Conversion experiences in lists

@@

ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
kable(df[df$price>=20,'name'])
```

@@

![The advent of recommender systems](figure/resnick_varian_page1_197.png)

@@

```{r arules, cache=TRUE, echo=TRUE, fig.margin=TRUE}
library(knitr)
library(arules)
library(arulesViz)
library(datasets)
data(Groceries)
itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))

@@

```{r tesco, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
    library(ggplot2)
    tesco = read.csv('data/reference_data/tesco_groceries.csv')
    total = sum(tesco$actual, na.rm=TRUE) 
    tesco = na.omit(tesco)
    tesco$actual = as.integer(tesco$actual)

@@

People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. [^45] But it also suggests a transformation in the way that lists exist in the world. They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.

@@

HERE

We have seen two different ways in which lists might be constructed in a recommender system. 
The first uses history and existence of previous lists, including the customer's own list, to populate a new list. 
The second uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
The 'relevancy model,' however, tries to do something different. 
The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.

@@

In some ways, Hansen's account resembles the Badiouean ontology of the multiple in which a closed set is surrounded and conditioned by an open set that is full of many more potentials. However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?

@@

It was surprising that Patel spoke relatively briefly about the new models. She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others. She dismissed most of the models quite quickly and focused only on one, the logistic regression model. Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.  The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.

@@

The graph compares models in terms of their predictions and how those predictions came to pass. The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new models (logistic regression, random forest, etc) for customers. An increase in precision means more sales.  Her interest lay in the relevance of the recommendations, and in particular, in how 'most of the gain comes from thinking of good features.'

The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems. The models she mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things. Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.

@@

Say Tesco has 1 million online customers. A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size. Given that any one customer is like to only have bought 100 different products, the data matrix will be `r format(100-(100/2e5)*100, 2)`% empty.

```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap ='The sparsity of data' }
library(ggplot2)
library(reshape2)
p = as.logical(rbinom(n=2e5, size=1, prob=0.001))
head(p, 30)
m = (matrix(p, nrow=1000, byrow=TRUE))

@@

So the raw data -- the 52 weeks of purchases for each customer -- can be aggregated to increase the density of the data in ways that helps the model. Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.

>>>>

---
title: "'Have you forgotten?' Conversion experiences in shopping lists"
author: "Adrian Mackenzie"
date: "`r Sys.Date()`"
output:
tufte::tufte_handout: default
citation_package: biber
latex_engine: xelatex
bibliography: ~/ref_bibs/uni.bib
link-citations: yes
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

# 'Have you forgotten?' Conversion experiences in shopping lists

@@

ggplot(df, aes(x=price,y=ContentsQuantity)) + geom_point() + theme(axis.text=element_text(size=10)) 
kable(head(df[df$price>=30,c('name','price')]))
```

@@

![The advent of recommender systems](figure/resnick_varian_1997_p1.png)

@@

```{r arules, cache=TRUE, echo=FALSE, fig.margin=TRUE}
library(knitr)
library(arules)
library(arulesViz)
library(datasets)
data(Groceries)
itemFrequencyPlot(Groceries,topN=10,col=rainbow(10), ntype="absolute",  cex.axis = 0.55)
rules <- apriori(Groceries, parameter = list(supp = 0.001, conf = 0.8))

@@

```{r tesco, echo=FALSE, cache=TRUE, fig.cap ='Tesco surplus',  fig.margin=TRUE}
    library(ggplot2)
    tesco = read.csv('data/reference_data/tesco_groceries.csv')
    total = sum(tesco$actual, na.rm=TRUE) 
    tesco = na.omit(tesco)
    tesco$actual = as.integer(tesco$actual)

@@

People inhabit lists and work with them no matter how they are made; thus vindicating ethnomethodologists assertion that social order is the product of ongoing sense-making practices. [^45] But the predicted list also suggests a transformation in the way that lists exist in the world. They are less the inscriptions of a discrete agent than an effect of an environment teeming with possible combinations and complementaries, a surplus of possible purchases that the supermarket needs to continually maintain but also filter in order to make it possible for something to happen.

@@

We have seen one ways in which lists might be constructed in a recommender system.[^56]

[^56]: The first uses history and existence of previous lists, including the customer's own list, to populate a new list.

It uses many shopping lists to identify associations between things and then uses the strongest associations to recommend additions to a current list. 
The 'relevancy model,' however, tries to do something different. 
The problem that the proliferation of things on the shelves of supermarket or grocery warehouse confronts is partly combinatorial.
With  `r total` products (actually Patel mentions 200,000 products) can be combined in truly astronomical numbers of combinations.
If a typical shopping list has 20 items, then there are `r format(choose(total, 20), 2)` possible combinations.
Some of this vast number of possible shopping lists is highly unlikely in terms of shopping lists and histories of shopping, but potentially highly interesting and value as recommendation.

@@

Hansen's account opposes the closed prediction to the open propensities of the data-world. A closed set is surrounded and conditioned by an open set that is full of many more potentials. However, if we think of the  new model that Patel describes as the delimited (or delimiting?) system within   the larger surplus of sensibility, how that help us understand what the 'Have you forgotten?' recommendations do?

@@

It was surprising that Patel spoke relatively briefly about the new models. She showed a graph that compared the old rule-based model with some of the new alternatives -- logistic regression, random forests, and a few others. She dismissed most of the models quite quickly and focused only on one, the logistic regression model. Or more accurately, she just said that logistic regression did better than other, and there was little more to discuss on that front.

@@

The graph compares models in terms of their predictions and how those predictions came to pass. The two columns on the left show earlier models -- the global model, which is not personalized, the association rules model for individual customers, and then the new relevance models (logistic regression, random forest, etc) for customers. An increase in precision means more sales.  Her interest lay in the relevance of the recommendations, and in particular, in how 'most of the gain comes from thinking of good features.'

The fact that so little was said about the models themselves suggests that they are now largely taken for granted in recommender systems.[^57] The models Patel mentioned are all classifiers, predictive models that treat a set of variables in a datasets as a vector space in which good classification means finding a way of defining a decision boundary between points that belong to different classes of things. Here the classes of things are binary: `recommended` or `not recommended.` Some classifier models, including logistic regression, generate a probability rather than simply a classification.

[^57]: The measure of doing better here is called precision, and it measures the proportion of the recommended products that are actually purchased.

So, the model generates probabilities for each product for each customer. The 'good features' strengthen associations between people and things. Each thing has a propensity in relation to each customer.

@@

Say Tesco has 1 million online customers. A data matrix for just whether they ever bought a particular product or not will be `r format(1e6 * 2e5)` in size. Given that any one customer is like to only have bought 100 different products, the probability matrix will be `r format(100-(100/2e5)*100, 2)`% empty.

```{r sparsity, echo=FALSE, cache=TRUE, fig.margin=TRUE, fig.cap ='The sparsity of data' }
library(ggplot2)
library(reshape2)
p = as.logical(rbinom(n=2e5, size=1, prob=0.001))
m = (matrix(p, nrow=1000, byrow=TRUE))

@@

So the raw data -- the 52 weeks of purchases for each customer -- can be rewoven with the database of Tesco products to increase the density of the data in ways that helps the model. Patel mentions some of the 'features' used in the new model but somewhat strangely at this point, her presentation turns toward the difficulties of measuring the success of the model and the infrastructural limitations on making recommendations in ways that are more alive to what is happening right now.
